---
title: "Module 11: Programming in R"
author: "Data-Scientific Programming: Lawrence University"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 6)
```

This module pertains to chapters 17 -21 in Wickham and Grolemund's text. We'll learn about basic programming in R. 

```{r}
library(tidyverse)
```

## 11.1: Vectors

Let's investigate how R performs operations on vectors. We'll start by generating four vectors. 

```{r}
x <- 1:10    # integers from 1 through 10
y <- seq(from = 2, to =29, by =3) # count from 2 to 29 by increments of 3
z <- rep(7, 5) # repeat 7, 5 times
w <- sample(1:10, 6) #randomly select six integers from 1 to 10
```

```{r}
x
```


```{r}
y
```


```{r}
z
```


```{r}
w
```


Perform each of the following operations in R. Write a sentence explaining what each operation does. If anything surprises you, note that. 

1) This tells you the data type of the variable you put in as a parameter.
```{r}
typeof(x)
```

2)This adds 5 to every element in the x array.
```{r}
x + 5
```

3) This adds each respective element of x and y together
```{r}
x + y
```

4) This multiplies every element of x and y together
```{r}
x * y
```

5) Similar as above, this adds the elements of x, and z. Z is recycled here.
```{r}
x + z
```

6) W is recyled and added to x here.
```{r}
x + w
```


7)This sorts x from smallest to greatest
```{r}
sort(x)
```

8) This gives the value of the length of the array
```{r}
length(x)
```

9)This gives the mean value of the array elements added together.
```{r}
mean(x)
```

10)This outputs a boolean to every variable whether or  not it is greater than 5.
```{r}
x > 5
```

11) This takes the mean of all values in the array that are greater than 5
```{r}
mean(x > 5)
```

12) This is taking mod 5 of every element and returning a boolean if it is a factor.
```{r}
x %% 5 == 0 
```

13)This selects the first element of the y array
```{r}
y[1]
```

14) This selects the third element of the y array
```{r}
y[3]
```

15) This selects multiple elements of y, here it is selecting the third and first element.
```{r}
y[(c(3,1))]
```

16) This displays the 1-5 elements of y.
```{r}
y[1:5]
```

17) This selects everything EXCEPT the second element of y
```{r}
y[-2]
```

18) This selects the elements that x mod 5 == 0 returns. (5, and 10)
```{r}
y[x %% 5 == 0]
```


## 11.2: Functions

A function takes in one or more inputs and returns one or more outputs. 

### Real Number Inputs

##### Example 11.2.1: 

Write a function called `sum_of_squares` that takes in two real numbers x and y, and returns the sum of the squares of these numbers. 

```{r}
sum_of_squares <- function(x, y) {
  return(x^2 + y^2)
}

sum_of_squares(2,3)
```


### Vectors and Dataframes as Inputs

The `starwars` dataset contains information on characters that appear in the Star Wars movies. 

```{r}
data(starwars)
str(starwars)
```


```{r}
ggplot(data=starwars, aes(x=mass)) + geom_histogram() + ggtitle("Mass Distribution")
```

```{r}
starwars %>% arrange(desc(mass))
```


```{r}
ggplot(data=starwars, aes(x=height)) + geom_histogram() + ggtitle("Height Distribution")
```

```{r}
starwars %>% arrange(desc(height))
```

```{r}
ggplot(data=starwars, aes(x=birth_year)) + geom_histogram() + ggtitle("Birth Year Distribution")
```

```{r}
starwars %>% arrange(desc(birth_year))
```


```{r}
mean(starwars$mass, na.rm=TRUE)
```

Why might this not be a good representation of the average mass of star wars characters?
We have a very broad dataset for mass. Starting at 15.0 and going all the way to 1358. Jabba is an obvious outlier and therefore is throwing the data as a whole off. 


A **trimmed mean** is a version of the mean that excludes the largest and smallest m numbers (where m is an integer). Calculate trimmed means for mass, height, and birth year that exclude the maximum and minumum values. 

Why is the following code sub-optimal

It's kind of clunky, we take the entire mass, height, and birth_year and subtract the max and the min from the data set. Then we divide by the entire sum of the data set. We have to do this for every variable we want a trimmed_mass off
```{r}
#mass calculation
(sum(starwars$mass, na.rm=TRUE) - min(starwars$mass, na.rm=TRUE) - max(starwars$mass, na.rm=TRUE)) / (sum(!is.na(starwars$mass)) - 2)
```

```{r}
#height calculation
(sum(starwars$height, na.rm=TRUE) - min(starwars$height, na.rm=TRUE) - max(starwars$height, na.rm=TRUE)) / (sum(!is.na(starwars$height)) - 2)
```

```{r}
#age calculation
(sum(starwars$birth_year, na.rm=TRUE) - min(starwars$birth_year, na.rm=TRUE) - max(starwars$birth_year, na.rm=TRUE)) / (sum(!is.na(starwars$birth_year)) - 2)
```


##### Example 11.2.2 

Write a function that calculates the trimmed mean, excluding the maximum and minimum value for a given vector. Exclude NA's from the calculation.

```{r}
vec <- starwars$mass
trimmed_mean <- function(vector){
  (sum(vector, na.rm=TRUE) - min(vector, na.rm=TRUE) - max(vector, na.rm=TRUE)) / (sum(!is.na(vector)) - 2)
}
trimmed_mean(vec)
```


#### Example 11.2.3

Write a function that inputs a dataframe and returns every fifth row. 

```{r}
every_fifth <- function(df){
  df.index<-as.numeric(rownames(df))
  df[df.index %% 5 == 0,]
}
fifth_output <- every_fifth(starwars)
```


#### Practice 11.2.4

Write functions to perform each of the following. Exclude NA's from the calculation. Test the functions on the starwars dataset, where appropriate. 

a) Input two real numbers x and y, and return the sum of the absolute values. 

```{r}
sum_absolute <- function(x, y) {
  if(x < 0){
    x = -x
  }
  if(y < 0){
    y = -y
  }
  return(x+y)
}
sum_absolute(-1,3)
```

b) Input two real numbers x and y and return the larger of the two. 

```{r}
compare <- function(x, y){
  if(x > y)
    return(x)
  return(y)
}

compare(5,3)
```

c) Input a vector and calculate the range of values (max - min). 

```{r}
calc_range <- function(vec) {
  return(max(vec, na.rm=TRUE) - min(vec, na.rm=TRUE))
}

calc_range(w)
```

d) Input a vector and calculate the variance. Note: $\frac{1}{n-1}\sum(x-\bar{x})^2$. Do not use the built-in function `var`.    

```{r}
calc_variance <- function(vec) {
  n <- length(vec)
  m <- mean(vec, na.rm = TRUE)
  sq_err <- (vec - m)^2
  sum(sq_err) / (n - 1)
}

calc_variance(w)
```

e) Input a dataframe and return 5 rows, selected at random. Hint: use `sample`. 

```{r}
select_rand <- function(df) {
  randomNums <- sample(1:NROW(df), 5)
  df[randomNums,]
}
randStarwars <- select_rand(starwars)
```


##### Example 11.2.5 

Write a function that takes in a function defined on the real numbers, and returns a graph, using a viewing window specified by the user. 

```{r}
plot_function <- function(n, f, xrange, yrange){
  # input f is a function on reals
  # input xrange is a vector for x-axis
  # input yrange is a vector for y-axis
  p <- ggplot(data = data.frame(xrange, yrange), aes(x=xrange, y=yrange)) + stat_function(fun=f) + xlim(xrange) + ylim(yrange)
return(p)
}
```

```{r}
square <- function(x){
  #input a real number
  #return its square
  return(x^2)
}

```


```{r}
plot_function(f=square, xrange = c(-2, 2), yrange = c(0,5) ) 
```

## Approximating Area

##### Example 11.2.6

The standard normal probability density function is given by $f(x)=\frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}$ We define the function in R. 

```{r}
normal_density <- function(x){
#input real number x
#return standard normal density at x   
  return(1/sqrt(2*pi)*exp(-x^2/2))
  }
```

Note: this function is actually already available via the built-in function `dnorm`

```{r}
plot_function(f=normal_density, xrange = c(-3, 3), yrange = c(0,0.5) ) 
```


This function does not have a closed-form anti-derivative, so we cannot find areas under the curve using standard calculus techniques. Instead, we can use simulation to approximate area. This technique is called **Monte Carlo Integration**. 

We'll use the `runif` function, which generates a specified number of random numbers (n), between a minimum (min) and maximum (max) value. 

```{r}
x<- runif(n=5, min=0, max=1)
y<- runif(n=5, min=0, max=1)

```


##### Describe the steps you could use to approximate the area under the standard normal curve between $x=-1$ and $x=1$, using simulation. 




##### Write a function that performs these steps. Allow the user to input the following arguments. Can you make the function return both an approximation of the area, and also a visual representation?

Arguments:     
n: number of points to simulate
f: function to integrate
x1: minimum of x range    
x2: maximim of x range      
y1: minimum of y range    
y2: maximum of y range    

```{r}
simulation <- function(num, fun, x1, x2, y1, y2){
  x<- runif(n=num, min=x1, max=x2)
  y<- runif(n=num, min=y1, max=y2l)
  plot_function(f=fun , xrange=c(x1, x2), yrange=c(y1,y2) )
}

simulation(3, square, -2, 2, 0, 5)
```


Test the function using n=10, n=1000, n=10000, and n=100000. 

```{r}

```



