---
title: "Module 9: Webscraping part 2"
author: "Data-Scientific Programming: Lawrence University"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 6)
```

```{r}
library(tidyverse)
library(rvest) #used to scrape data from web
library(robotstxt) #used to check whether a website can be scraped
```


## Section 9.2 Second Example

Go to the [link](https://www.imdb.com/search/title/?title_type=feature&release_date=2019-01-01,2019-12-31&sort=boxoffice_gross_us,desc). 

This site contains information on the 50 highest-grossing movies of 2019 (according to US box office revenue). We'll scrape this information. 

#### 1) Read in page

Read in the url

```{r}
page <- read_html("https://www.imdb.com/search/title/?title_type=feature&release_date=2019-01-01,2019-12-31&sort=boxoffice_gross_us,desc")
```

#### 2) Scrape titles

Use Selector Gadget to help read in the titles. Format them as text. 

```{r}
title <- page %>%
  html_nodes(".lister-item-header a") %>%
  html_text()
```

#### 3) Scrape ratings

Use Selector Gadget to help read in the ratings. Be sure to only scrape the ratings information. Format them as numeric. 

```{r}
rating <- page %>%
  html_nodes(".ratings-imdb-rating strong") %>%
  html_text() %>%
  as.numeric()
```

#### 4) Scrape MPAA ratings

Read in the MPAA ratings (G, PG, PG-13, R, NC-17). Format them as character. 

```{r}
MPAA_rating <- page %>%
  html_nodes(".certificate") %>%
  html_text() %>%
  as.character()
```

What is the length of this vector? Why does it not match the lengths of the title and ratings vectors?

```{r}
length(title)
```

```{r}
length(MPAA_rating)
```

There are 50 titles, but only 49 ratings. Notice that the film in line 37 does not have a rating. This will create problems when we try to put these in a tibble. Later on, we'll find a better way to scrape MPAA_rating. 

#### 5) Scrape runtimes

Read in the runtimes. We'll use `str_remove_all()` to eliminate excess text. Format as numeric

```{r}
runtime <- page %>%
  html_nodes(".runtime") %>%
  html_text() %>%
  as.character() %>% 
  str_remove_all(" min") %>% 
  as.numeric()
```

#### 6) Scrape Genre

Scrape movie genres. Eliminate the "\n" and get rid of excess spacing. Format as character. 

```{r}
genre <- page %>%
  html_nodes(".genre") %>%
  html_text() %>%
  as.character() %>%
  str_remove_all("\n") %>%
  trimws()
```

####  7) Scrape Metascore

Scrape metascores. Eliminate the "\n" and get rid of excess spacing. Format as numeric. 

```{r}
metascore <- page %>%
  html_nodes(".metascore") %>%
  html_text() %>%
  as.character() %>%
  str_remove_all("\n") %>%
  str_trim() %>% 
  as.numeric()
```

####  8) Scrape Votes

Scrape number of votes. Remove all commas and convert to numeric. 

```{r}
votes <- page %>%
  html_nodes(".sort-num_votes-visible span:nth-child(2)") %>%
  html_text() %>%
  str_remove_all(",") %>%              # remove all commas in each entry
  as.numeric() %>% as.numeric()
```

#### 9) Scrape Gross

Scrape film gross. Remove excess letters and characters and convert to numeric. 

```{r}
gross <- page %>%
  html_nodes(".ghost~ .text-muted+ span") %>%
  html_text() %>%
  str_remove_all("\\$") %>%  #note \\ is required before special characters
  str_remove_all("M") %>%
  as.numeric()
```
  

#### 10) Better way to scrape MPAA and metascore

scrape MPAA rating and metascore in a way that results in vectors of length 50, with NA's for missing value(s). 

```{r}
MPAA_rating <- page %>%
  html_nodes(".lister-item-header+.text-muted") %>%
  html_text() %>%
  str_replace("\n            \n", "\n           NA \n |" ) %>%
  str_split("\\|") %>%
  map(1, .default=NA)%>%
  trimws() %>%
  as.character()
  head(MPAA_rating)
```

```{r}
metascore <- page %>%
  html_nodes(".ratings-bar") %>%
  html_text() %>%
  str_remove_all(" ") %>%
  str_remove_all("\n") %>%
  str_split("X") %>%
  map(2, .default=NA) %>%
  str_remove_all("Metascore") %>%
  str_trim() %>%
  as.numeric()
head(metascore)
```

